---
title: "Aides Drug"
author: "Isabela Bercian"
date: "18 March 2025"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

First, we need to take the data that's in triUmp and produce a subset consisting of only the females. 

# Results

```{r}
females <- filter(triUmp, gender=="F")
```


<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment, data=females)
```



```{r}
barchartGC(~infected+treatment, data=females, type = "percent")
```

These graphs show that the drug in question is not effective in accomplishing its intended purpose. 


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1<- xtabs(~infected+treatment, data=triUmp)
rowPerc(table1)
colPerc(table1)
```

These are numerical results which show that, of those infected with AIDS, 59 percent of them had taken the drug, and 40 percent of them had taken the placebo. This means that most of the people who retracted AIDS had taken the drug which was meant to prevent it. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Because the results are such a low number in these statistical tests, we will reject the null hypothesis. The first test gave a p value of .00000002761. The second gave 0. In the last test, the p value was .000000233. These are all very low p values. In all of these cases, the null hypoy=thesis must be rejected. The odds ratio goes from .63 to .81. This means that the odds of getting this desease is 0.71. This means there are high chances of getting AIDS if you take this drug. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

According to all statistical evidence, the drug in question does not achieve its intended purpose. It does not work.
